An Introduction to Rocker:\\ Docker Containers for R

### 2017-11-24 boettiger-eddelbuettel.Rmd
whitespace mostly

To make extensions transparent and persistent, Rocker images can be extended by
any user by writing their own Dockerfiles based on an appropriate Rocker image.
The Dockerfiles in the Rocker stack should themselves provide a simple example
of this, (as described in the following section).  
A user begins by selecting an appropriate base image for their needs: 
if the RStudio\textsuperscript{\textregistered} interface is desired, a user might start with `FROM rocker/rstudio`;
an image for testing an R package with compiled code might use `FROM rocker/r-devel-san`,
and an image for reproducing a data analysis will probably select a stable version tag
in addition to an appropriate base library, _e.g._,: `FROM rocker/tidyverse:3.4.1`
Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. Details on how to extend Rocker images can
be found at <https://rocker-project.org>.

@@

As the names `testing` and `unstable` imply, particular versions of package can
change as packages move from `unstable` into `testing`. New versions
are sent to `unstable` during the normal course of Debian development.
This can occasionally break an previously-working installation command
in a Dockerfile until the maintainer redirects the package manager to
install a package from the `unstable` sources that could previously
be installed from `testing`, or vice versa (using the `-t` option in
`apt`).  That said, packages only migrate from 
`unstable` to `testing` after a period of several days---and if the 
migration and installation of the particular version is free of 
interactions with other packages in their dependency graph. That way,
`unstable` serves as validation lab which leaves `testing` reasonably 
stable yet current.

>>>>

To make extensions transparent and persistent, Rocker images can be extended by
any user by writing their own Dockerfiles based on an appropriate Rocker image.
The Dockerfiles in the Rocker stack should themselves provide a simple example
of this, (as described in the following section).
A user begins by selecting an appropriate base image for their needs: 
if the RStudio\textsuperscript{\textregistered} interface is desired, a user might start with `FROM rocker/rstudio`;
an image for testing an R package with compiled code might use `FROM rocker/r-devel-san`,
and an image for reproducing a data analysis will probably select a stable version tag
in addition to an appropriate base library, _e.g._,: `FROM rocker/tidyverse:3.4.1`.
Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. Details on how to extend Rocker images can
be found at <https://rocker-project.org>.

@@

As the names `testing` and `unstable` imply, particular versions of package can
change as packages move from `unstable` into `testing`. New versions
are sent to `unstable` during the normal course of Debian development.
This can occasionally break a previously-working installation command
in a Dockerfile until the maintainer redirects the package manager to
install a package from the `unstable` sources that could previously
be installed from `testing`, or vice versa (using the `-t` option in
`apt`).  That said, packages only migrate from 
`unstable` to `testing` after a period of several days---and if the 
migration and installation of the particular version is free of 
interactions with other packages in their dependency graph. That way,
`unstable` serves as validation lab which leaves `testing` reasonably 
stable yet current.

### 2017-11-24 boettiger-eddelbuettel.tex
whitespace mostly

Docker is a popular open-source tool to create, distribute, deploy, and
run software applications using \emph{containers}. Containers provide a
virtual environment (see \citet{Clark2014} for an overview of common
virtual environments) requiring all operating-system components an
application needs to run. Docker containers are lightweight as they
share the operating system kernel, start instantly using a layered
filesystem which minimizes disk footprint and download time, are built
on open standards that run on all major platforms (Linux, Mac, Windows),
and provide an added layer of security by running an application in an
isolated environment \citep{what-docker}. Familiarity with a few key
terms is helpful in understanding this paper. The term ``container''
refers to an isolated software environment on a computer. R users can
think of running a container as analogous to loading an R package; a
container is an active instance of a static Docker image. A Docker
``image'' is a binary archive of that software, analogous to an R binary
package: a given version is downloaded only once, and can then be
``run'' to create a container whenever it is needed. A ``Dockerfile'' is
a recipe, the source-code, to create a Docker image.\\
Pre-built Docker images are publicly available through Docker Hub, which
plays a role for central distribution similar to CRAN in our analogy.
Development and contributions to the Rocker project focus on the
construction, organization and maintenance of these Dockerfiles.

@@

This sandboxing feature is also valuable in the remote computing
context, allowing system administrators to grant users freedom to
install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access is
required to launch Docker containers, though not to access containers
already running and providing some service such as RStudio. Users
logging into a container through the
RStudio\textsuperscript{\textregistered} interface do not by default
have root privileges, though are able to install R packages. Granting
these users root privileges in the container still leaves them sandboxed
from the host container.\\
Sandboxing also serves an important function in reproducible research by
making it easier to test a specified environment in isolation from the
host machine. Unlike traditional virtual machines, containers do not
impose a large footprint of reserved resources as a typical host can
easily support 100s of containers \citep{what-docker}.

>>>>

Docker is a popular open-source tool to create, distribute, deploy, and
run software applications using \emph{containers}. Containers provide a
virtual environment (see \citet{Clark2014} for an overview of common
virtual environments) requiring all operating-system components an
application needs to run. Docker containers are lightweight as they
share the operating system kernel, start instantly using a layered
filesystem which minimizes disk footprint and download time, are built
on open standards that run on all major platforms (Linux, Mac, Windows),
and provide an added layer of security by running an application in an
isolated environment \citep{what-docker}. Familiarity with a few key
terms is helpful in understanding this paper. The term ``container''
refers to an isolated software environment on a computer. R users can
think of running a container as analogous to loading an R package; a
container is an active instance of a static Docker image. A Docker
``image'' is a binary archive of that software, analogous to an R binary
package: a given version is downloaded only once, and can then be
``run'' to create a container whenever it is needed. A ``Dockerfile'' is
a recipe, the source-code, to create a Docker image. Pre-built Docker
images are publicly available through Docker Hub, which plays a role for
central distribution similar to CRAN in our analogy. Development and
contributions to the Rocker project focus on the construction,
organization and maintenance of these Dockerfiles.

@@

This sandboxing feature is also valuable in the remote computing
context, allowing system administrators to grant users freedom to
install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access is
required to launch Docker containers, though not to access containers
already running and providing some service such as RStudio. Users
logging into a container through the
RStudio\textsuperscript{\textregistered} interface do not by default
have root privileges, though are able to install R packages. Granting
these users root privileges in the container still leaves them sandboxed
from the host container. Sandboxing also serves an important function in
reproducible research by making it easier to test a specified
environment in isolation from the host machine. Unlike traditional
virtual machines, containers do not impose a large footprint of reserved
resources as a typical host can easily support 100s of containers
\citep{what-docker}.
